### 2016-12-29 paper_drafts/RJournalRevisions/anderson-eason-barnes.Rmd

---
title: >
    Working with daily climate model output data in R and the \pkg{futureheatwaves} package
author:
  - name: G. Brooke Anderson
    affiliation: Colorado State University
    address:
    - Department of Environmental & Radiological Health Sciences
    - 1681 Campus Delivery
    - Fort Collins, Colorado 80523
    email:  brooke.anderson@colostate.edu
  - name: Colin Eason
    affiliation: Colorado State University
    address:
    - Department of Computer Science
    - 1873 Campus Delivery
    - Fort Collins, Colorado 80523
    email:  aimesce@gmail.com
  - name: Elizabeth A. Barnes
    affiliation: Colorado State University
    address:
    - Department of Atmospheric Science
    - 1371 Campus Delivery
    - Fort Collins, CO 80523
    email:  eabarnes@atmos.colostate.edu
abstract: >
  Research on climate change impacts can require extensive processing, to incorporate output from ensembles of different climate models and different simulations of each model. This processing can be particularly extensive when identifying and characterizing extreme events like heat waves and frost day spells, as these must be processed from model output with daily time steps and may use identification threholds based on local distributions (e.g., 98th percentile temperature for that location). Here, we provide an overview of working with daily climate model output data in R. Further, we present the \pkg{futureheatwaves} package, which we developed to ease the process of identifying, characterizing, and exploring multi-day extreme events in climate model output. This package takes as input a directory of projection files from many climate models, identify all heat waves for multiple cities using a user-specified heat wave definition, and write an output directory of files with these heat waves and their characteristics (e.g., length, absolute and relative temperatures). The \pkg{futureheatwaves} package then allows the user to create custom functions and apply them across all heat wave files; this functionality can be used either to explore heat wave patterns (e.g., average length and intensity across different climate models) or to apply epidemiological estimates to project health impacts under different scenarios.
preamble: >
  % Supposedly allows you to add stuff to LaTeX preamble. Not for me.
output: rticles::rjournal_article
---

```{r echo = FALSE, message = FALSE, warning = FALSE}
library(futureheatwaves)
library(ggplot2)
```

# Introduction

Research on climate change impacts can require extensive processing. This is not only because output files for a single climate model can be large, but also because of the rising popularity of ensemble techniques \citep{IPCCch1}, in which, to better characterize uncertainty in projections, impacts are assessed for multiple climate models, multiple simulations of each climate model, and multiple climate experiments. Projections of regional climate change over the next century are subject to high uncertainty due to three distinct sources: (1) internal climate variability, i.e. climate noise, (2) climate model uncertainty, i.e. the same forcing can produce a different response in different models and (3) scenario uncertainty, i.e. uncertainty in future climate forcings (e.g. \citet{hawkins2009potential}). One approach for ensuring that these sources of uncertainty are characterized is to simulate the future climate many times with multiple models and for multiple future scenarios.

For example, the Coupled Model Intercomparison Project, phase 5 (CMIP5; \citet{taylor2012overview}) brought together dozens of major climate modeling groups around the world to simulate the same future radiative forcing scenarios, but with their own models. This created an ensemble of state-of-the-art climate model projections that allow researchers to study projections and their uncertainties. Most of these modeling groups additionally performed more than one simulation for each scenario and model (i.e. multiple ensemble members), perturbing the initial conditions by a very tiny amount to quantify uncertainties due to internal climate variability.

This processing is particularly intensive for smaller time steps, like daily climate model output. While some climate impacts can be assessed using climate model output at monthly, seasonal, or yearly time steps, the impacts of multi-day extreme events must be assessed using output in daily time steps. Such multi-day extreme events include heat waves, cold spells, frost day spells, and droughts. The assessment is further complicated if extreme events are identified based on conditions that are rare for a certain location (e.g., 98th percentile of local temperature distribution for identifying heat waves) \citep{IPCCch1}, as this event definition must be determined at each study location from climate model output. Further, it is often of interest to create summaries of multiple characteristics of these extreme events. One interest, for example, in interpreting climate change scenarios output is whether frequency of heat waves or warm spells will change, as well as how certain characteristics of these extreme events (e.g., length, intensity) will change under different scenarios \citep{IPCCch1}.

We begin this article with an overview of climate model output data, particularly daily data, for R users. We outline where data from CMIP5 can be obtained as well as how to work with the file format (netCDF) from R. We overview some R packages that can be useful when working with this data, as well as aspects of the data (e.g., non-standard calendars) of which users should be aware when working with daily climate model output in R.

After this overview, we present the \pkg{futureheatwaves} package. which we created to aid in identifying and characterizing any type of multi-day extreme event from daily climate model output (Table \ref{tab:goals}). Further, this package provides some functionality particularly useful in identifying and characterizing heat waves specifically. Quantifying the impacts of heat waves on human health suffer from additional sources of uncertainty beyond those inherent in projections of regional changes in surface temperature. These include: (1) uncertainty in the ability of communities to adapt to changing temperatures (the adaptation scenario) and (2) uncertainty in the definition of a heat wave itself. Thus, identifying and characterizing the impacts of future heat waves in state-of-the-art climate models requires analyzing hundreds of projections based on combinations of anthropogenic activity, climate model and ensemble member, adaptation scenario, and heat wave definition. Such an analysis is non-trivial. For example, to estimate the possible impacts of heat waves using 32 ensemble members of CMIP5, two forcing scenarios, two different heat wave definitions, and five progressive assumptions about adaptation to heat, would require that one identify and characterize heat waves in 640 separate temperature time series for each region of interest.

\begin{table}[t]
\begin{center}
\newcolumntype{L}[1]{>{\raggedright\arraybackslash}p{#1}}
\begin{tabular}{lL{\dimexpr0.8\textwidth-2\tabcolsep\relax}}
\toprule
 & Design goals of the \pkg{futureheatwaves} package \\
\midrule
1 & Make processing of large sets of climate projections more practical for researchers exploring the potential impacts of heat waves and other multi-day extreme events.\\
2 & Speed up processing time by incorporating C++ in event identification.\\
3 & Keep track of the names of climate models, and number of ensemble members processed for each.\\
4 & Not only identify, but also characterize, all extreme events within each climate projection, to allow the exploration of patterns in these characteristics across different projections and also to allow the use of more complex impacts models, including models that include effect modification by event characteristics (e.g., event length, event intensity). For example, this package allows the user to apply a health effects model where risk of mortality is not the same for every heat wave, but rather is modified by heat wave length, intensity, or other measured characteristics.\\
5 & Give users extensive power in customizing the process, including allowing custom event definitions.\\
6 & Allow users to easily explore the extreme events identified within all climate projections by applying custom functions across heat wave data sets from all projections at once.\\
7 & Create output that is in a ``tidy" data format, allowing it to work well with \pkg{ggplot2} for visualization.\\
\bottomrule
\end{tabular}
\end{center}
\caption{Design goals for the \pkg{futureheatwaves} package.}
\label{tab:goals}
\end{table}

The \pkg{futureheatwaves} package automates the process of identifying and characterizing multi-day extreme events across different ensemble members of one or more climate models. A variety of different heat wave definitions have been used to identify heat waves in a time series of temperature data \citep{smith2013heat}, and the choice of heat wave definitions can influence both health effect estimates \citep{chen2015influence, kent2014heat} and projected heat wave trends \citep{smith2013heat}. Further, other types of extreme events will be defined differently than heat waves (for example, frost day spells may be defined as one or more days with temperature at or below $0^oC$). This package therefore allows the user to create and use a custom the extreme event definition used to identify events in the climate model output. Finally, this package allows the user to explore the extreme events identified in the climate model output with a function that can take a user-defined R function and apply it across all extreme event files generated for the separate ensemble members.

# An overview of climate model output for R users

## CMIP5 climate model output data

For climate impact studies, a top source for climate model output files is the Coupled Model Intercomparison Project, which is currently in its fifth phase (CMIP5). Over 20 climate modeling groups have created one or more climate models which, for this project, are run using standardized scenarios \citep{taylor2012overview}. The resulting output is uniform across modeling groups and has a consistent structure, which allows comparison of simulations from different models \citep{IPCCch9}. For CMIP5, each group created simulations under several experiments, with experiments varying in terms of the radiative forcing. This radiative forcing depends on time-varying model inputs (greenhouse gas emissions or concentrations, land use changes, etc.), which are specified for each experiment \citep{taylor2012overview, IPCCch9}. Experiments include historical experiments (run using forcings consistent with observed and reconstructed data for 1850--2005), pre-Industrial control experiments, and experiments of future scenarios of radiative forcing over the 21st century or longer (e.g., RCP4.5, RCP8.5) \citep{taylor2012overview}. Some modeling groups created ensembles of output for a specific model and experiment, in which they ran the experiment multiple times with the model with very small changes to the initial conditions, resulting in multiple ensemble members for a single climate model and experiment. CMIP5 climate model output is created at a number of different time steps (e.g., daily, monthly, seasonal, yearly) \citep{taylor2010cmip5}, and some variables are reported at multiple depths in the ocean or atmosphere (e.g., ocean temperature). Here, we will focus on data with a daily time step for variables reported at a single depth (e.g., near-surface air temperature).

The CMIP5 climate model output data is distributed across data nodes at different climate modeling centers \citep{taylor2012overview}, but can be accessed centrally at the World Climate Research Programme CMIP5 data portal at https://pcmdi.llnl.gov/search/cmip5/. Users must register before downloading data, and some data are restricted to non-commercial use. There is a separate file for each combination of climate model, experiment, modeling realm (e.g., atmosphere, ocean), variable, time step, and ensemble member \citep{taylor2012overview, taylor2010cmip5}. For finer time scales, this output is further split across multiple files for specific year ranges (e.g., 5 years of output for each file) \citep{taylor2010cmip5}. Filenames for CMIP5 files can be parsed to generate information about the output variable, climate model, experiment, and ensemble member for the simulation \citep{taylor2010cmip5}. Files can be searched and downloaded through a point-and-click web interface. The data portal also allows you to generate a specific wget script, which you can use to download many files at once. A wget script can also be generated through Earth System Grid Federation's Search RESTful API. Tips on efficiently searching and downloading the data, including through use of wget scripts and the search API, are available as user tutorials through the website of the University of Colorado Boulder's Earth System CoG (e.g., https://www.earthsystemcog.org/projects/cog/doc/wget for a tutorial on downloading files using wget).

CMIP5 files are saved in Network Common Data Format (netCDF). NetCDF is a binary file format that allows storage of data representing a regular array. These climate model output files can be as large as several gigabytes [?]. For climate model output at a single depth (e.g., near-surface air temperature), the data is a 3-dimensional array, with dimensions representing time and two metrics of location (e.g., latitude and longitude) (Figure \ref{fig:netcdfexample}). Global climate models generate output at regularly-spaced time steps, typically at regularly-spaced grid points around the world. The latitude and longitude spacing of grid points vary by climate model, but are typically 1--2 degrees for atmospheric variables in CMIP5 models \citep{IPCCch9}. Each data point in the netCDF array gives the modeled value of the variable (e.g., surface temperature) for a single time point and location. For CMIP5 climate model output, the location units are in degrees east and degrees north for longitude and latitude, respectively. For daily output files, the time unit is in days since a specified origin date-time (e.g., days since 1850-01-01 00:00:00) \citep{taylor2010cmip5}. All CMIP5 output files are requred to include certain metadata  \citep{taylor2010cmip5}. This required metadata includes the experiment, forcing agents input to the model to create the simulation, time step, institution and institutional contact information, climate model, and modeling realm \citep{taylor2010cmip5}. The metadata also must include units for all of the coordinate variables (e.g., longitude, latitude, time).

\begin{figure}
\begin{center}
\includegraphics[width = 0.9\textwidth]{netcdf_structure_example}
\end{center}
\caption{Example of structure of a NetCDF climate model output file for a variable reported at a single depth, like surface air temperature. Data are stored in a three-dimensional array, with measurements at each time step and grid location. Surface temperature data are typically indexed in climate model output files by longitude, latitude, and time, in that order. For example, if the air surface temperature ("tas") is read into an R object called \code{tas}, you can access the value for the first day at the fourth longitude and third latitude with \code{tas[4, 3, 1]}. In addition to the output variable (temperature in this example), vectors with the ordered values of each dimension (longitude, latitude, and time) can also be read in from the netCDF file, as well as attribute data (e.g., units for variables, the calendar used for time).}
\label{fig:netcdfexample}
\end{figure}

\citet{taylor2012overview} and \citet{meehl2007wcrp} are excellent resources for finding out more about the CMIP climate model output data.

## Working with climate model output in R

A few R packages can be used to work with the netCDF file format used for CMIP5 files. Earlier packages to work with netCDF files included \pkg{ncdf} and \pkg{ncvar}, but these do not work with the newer netCDF version 4 released in 2008 and are no longer available through CRAN. More recent packages, including \pkg{ncdf4} \citep{ncdf4} and \pkg{RNetCDF} \citep{michna2013rnetcdf, RNetCDF}, work with both version 4 and netCDF's older version 3. Climate model output data for CMIP5 is required to conform with the earlier version (version 3) \citep{taylor2010cmip5} and so should work with any of these packages, although it is safer to write code using functions that can be used with version for in case future phases of CMIP do not require files to conform with netCDF version 3.

The netCDF format allows you to access metadata and variables describing the dimensions of the data without reading the full file into memory. The metadata describes the dimensions of the array and the units, etc., of each variable and is printed by the print method of the R object returned by the \code{nc\_open} function of \pkg{ncdf4}. For CMIP5 files, the dimension variables include the times and locations corresponding to each array in the data. Before reading in a variable from a netCDF file, a connection to the file must be opened, for example with the \code{nc\_open} function from \pkg{ncdf4}. Variables can then be read in using the \code{ncvar\_get} function from \pkg{ncdf4}, with the \code{varid} parameter set to "lat", "lon", or "time" (as a caveat, many climate models output to non-Gregorian calendars, in which case the time variable should be read in using a different function, as discussed later in this section). The climate output variable (e.g., near-surface air temperature) can similarly be read in using \code{ncvar\_get}. In this case, the \code{varid} parameter should be set using the appropriate CMIP5 variable name (e.g., "tas" for near-surface air temperature); these variable names can be found in the CMIP Requested Output tables \citep{taylor2010cmip5}. In practice, you can use the dimensional time and location data to identify the location of the variable data you need in the netCDF array and use indexing to read only that data into memory, without needing to read in the full file \citep{michna2013rnetcdf}. Once the user is done reading in data from the file, the connection can be closed (e.g., with the \code{nc\_close} function from \pkg{ncdf4}).

Since the late 1500s, Western dates have been set using the Gregorian calendar, which has 365.2425-day years. Some climate models, however, output to different calendars, including the Julian calendar (365.25-day years), a calendar where there are no leap years (365-day years), a calendar where every year is a leap year (366-day years), and a calendar of twelve 30-day months (360-day years) \citep{cfconventions}. With these non-Gregorian calendars, R's base functions for converting a vector to a Date class based on the number of days since an origin date (\code{as.Date}, \code{as.POSIXct}) do not return the desired values. The \pkg{PCICt} \citep{PCICt} and \pkg{ncdf4.helpers} \citep{ncdf4.helpers} packages provide further functionality with netCDF files that can be useful when working with climate model output data and provide particular help in working with different calendars. CMIP5 netCDF files include information on the calendar used, which the \code{nc.get.time.series} function in \pkg{ncdf4.helpers} use to convert the "time" variable in the file to an object of the PCICt class, which provides Date-like functionality for 360- and 365-day calendars \citep{PCICt}. As a note, while these functions will help with handling most CMIP5 files, the CMIP5 standards allows use of other calendars which may not be succesfully handled by these functions, so it is important to assess whether the time variable range in the PCICt object correctly matches the expected date ranges for a file as you process CMIP5 data in R.

The size of the climate model output files can be large enough that it may make more sense to work with smaller chunks of the data, rather than reading all data into memory and working with the data all at once \citep{RCMIP5}. This problem aggregates when working with multiple climate models and more than one ensemble member for each of those climate models.

The following code gives an example of using a CMIP5 file and some of the R packages and functions discussed. Results include a map of near-surface air temperatures from a single climate model on a specific date (July 1, 2075) for China (Figure [x]) and a time series of daily near-surface air temperature simulations for the climate model grid point closest to Beijing, China (Figure [x]). To try this example, you will need to download the file ... from ... and save it in the "tmp" subdirectory of your home directory.

```{r}

```

In addition to these general packages for working with netCDF files, there are several R packages specifically for working with climate model output data, including \pkg{RCMIP5} \citep{RCMIP5} and \pkg{wux} \citep{wux}. However, these packages are more useful for working with data output at time steps of a month or higher and have limited utility with the daily climate model output data required for studies of multi-day extreme events.

The \pkg{RCMIP5} package includes functions to read in CMIP5 data from netCDF files, scan a directory of CMIP5 files and determine models with continuous available data, create objects of a special "cmip5data" class to work with CMIP5 data within R, and parse the filenames for all files in a directory to extract information within the filename. For this package, most functions only work with monthly or less frequent data (e.g., the functions \code{checkTimePeriod} and \code{cmip5data}, as well as functions that work with \code{cmip5data} objects including \code{filterDimensions} and \code{getProjectionMatrix}) \citep{RCMIP5}. While the \code{loadCMIP5} function does successfully load the daily data as a \code{cmip5data} object, most of the methods for this object type do not do anything meaningful for daily data. The package's \code{getFileInfo} function, however, will work with CMIP5 files of any time step; this function identifies all CMIP5 files in a directory and creates a dataframe that parses the information contained in the file name. As a note, the \code{get.split.filename.cmip5} function in the \pkg{ncdf4.helper} package similarly can be used to parse information contained in CMIP5 file names \citep{ncdf4.helpers}.

The \pkg{wux} package \citep{wux} includes functions that allow the user to download CMIP5 monthly-aggregated output directly from within R with the \code{CMIP5fromESGF} function. However, this function does not allow downloading of climate model output with finer time steps, like daily data. This package uses the \code{models2wux} function to read in climate model output netCDF files and convert to a WUX dataframe that can be used by other functions in the package. While this function can input climate model output with daily time steps, if the element "what.timesteps" of the \code{modelinput} list input is set to "daily", the function aggregates this data to a monthly or less frequent (e.g., seasonal) aggregation when creating the WUX dataframe. Therefore, while this package provides very useful functionality for working with averaged output of daily climate model output data, it cannot easily be used to identify and characterize multi-day extreme events like heat waves.

More tips on working with CMIP5 files in R are provided in the "starting_from_netcdf" vignette of the \pkg{futureheatwaves} package.

# The \pkg{futureheatwaves} package

## How the package works

\begin{widefigure}
\includegraphics[width = 0.9\textwidth]{OverviewFigure}
\caption{Overview of the functionality of the \pkg{futureheatwaves} package. The package takes a directory with climate projection files (left), for one or more climate models, with one or more ensemble members for each climate model (this example figure shows four climate models with one or two ensemble members each). The gen\_hw\_set function processes these files to create a data frame for each ensemble member, identifying and characterizing all multi-day extreme events (e.g., heat waves) in the time series projection for that ensemble member. The apply\_all\_models function allows users to explore these extreme events by applying user-created functions across all the extreme event data frames, creating a summary data frame with results.}
\label{fig:overview}
\end{widefigure}

The \pkg{futureheatwaves} package was facilitated to aid in identifying, characterizing, and exploring multi-day extreme events in daily climate model output data. Figure \ref{fig:overview} gives an overview of the two primary functions of the \pkg{futureheatwaves} package. First, the \code{gen\_hw\_set} function processes a directory of climate projection files that are stored locally on the user's computer (Figure \ref{fig:overview}, "Climate projections"), to generate a list of all extreme events in each projection, as well as over a dozen characteristics of each identified extreme event (Table \ref{tab:hwcharacteristics}). The extreme events are identified and characterized at one or more study locations (e.g., cities), with study locations specified by the user in an input file. The extreme events identified for each ensemble member are output as separate files in a directory specified by the user (Figure \ref{fig:overview}, ``Extreme events datasets"). Second, the \code{apply\_all\_models} function allows the user to apply custom functions across all the extreme event data frames generated by \code{gen\_hw\_set} to summaries of extreme events across all climate models and ensemble members (Figure \ref{fig:overview}, right).

\begin{table}
\newcolumntype{L}[1]{>{\raggedright\arraybackslash}p{#1}}
\begin{tabular}{lL{\dimexpr0.8\textwidth-2\tabcolsep\relax}}
\toprule
Column name & Description of characteristic \\
\midrule
\code{mean.var} & Average daily value of the variable across all days in the extreme event, in the units in which the variable is expressed in input files (e.g., average daily mean temperature during the heat wave in degrees Kelvin) \\
\code{max.var} & Highest daily value of the variable across all days in the extreme event, in the units in which the variable is expressed in input files \\
\code{min.var} & Lowest daily value of the variable across all days in the extreme event, in the units in which the variable is expressed in input files \\
\code{length} & Number of days in the event \\
\code{start.date} & Date of the first day of the event \\
\code{end.date} & Date of the last day of the event \\
\code{start.doy} & Day of the year of the first day of the event (1 = Jan. 1, etc.)\\
\code{start.month} & Month in which the event started (1 = January) \\
\code{days.above.abs.thresh.1} & Number of days in the event above a specified absolute threshold (default is the number of days in the event above $80^oF$ / $26.7^oC$, but this can be changed with the \code{absolute\_thresholds} argument in \code{gen\_hw\_set}) \\
\code{days.above.abs.thresh.2} & Number of days in the event above a specified absolute threshold (default is the number of days in the event above $85^oF$ / $29.4^oC$) \\ 
\code{days.above.abs.thresh.3} & Number of days in the event above a specified absolute threshold (default is the number of days in the event above $90^oF$ / $32.3^oC$) \\
\code{days.above.abs.thresh.4} & Number of days in the event above a specified absolute threshold (default is the number of days in the event above $95^oF$ / $35.0^oC$) \\
\code{days.above.99th} & Number of days in the event above the $99^{th}$ percentile of the variable for the location, using the period specified with the \code{referenceBoundaries} argument in \code{gen\_hw\_set} as a reference for determining these percentiles \\
\code{days.above.99.5th} & Number of days in the event above the $99.5^{th}$ percentile of the variable for the location, using the period specified with the \code{referenceBoundaries} argument in \code{gen\_hw\_set} as a reference for determining these percentiles \\
\code{first.in.year} & Whether the event was the first to occur in its calendar year in the location \\
\code{mean.var.quantile} & The percentile of the average variable value during the event compared to the location's year-round distribution of the variable, based on the variable distribution for the location during the period specified by the \code{referenceBoundaries} argument in \code{gen\_hw\_set} \\
\code{max.var.quantile} & The percentile of the maximum variable value during the event compared to the location's year-round distribution of the variable, based on the variable distribution for the location during the period specified by the \code{referenceBoundaries} argument in \code{gen\_hw\_set} \\
\code{min.var.quantile} & The percentile of the minimum variable value during the event compared to the location's year-round distribution of the variable, based on the variable distribution for the location during the period specified by the \code{referenceBoundaries} argument in \code{gen\_hw\_set} \\
\code{mean.seasonal.var} & The location's average seasonal value of the variable (by default, season is set to May--September, but this can be changed with the \code{seasonal\_months} argument in \code{gen\_hw\_set}), based on the variable values for the location during the years specified by the \code{referenceBoundaries} argument in \code{gen\_hw\_set} \\
\code{mean.yearround.var} & The location's average year-round value of the variable, based on the variable values for the location during the years specified by the \code{referenceBoundaries} argument in \code{gen\_hw\_set} \\
\bottomrule
\end{tabular}
\caption{Extreme event characteristics measured by the \code{gen\_hw\_set} function in the \pkg{futureheatwaves} package. The left column gives the name of each variable's column in the extreme event datasets created by the \code{gen\_hw\_set} function. When characterizing extreme events below a threshold, like cold spells, appropriate alternatives are given for some columns (e.g., \code{days.below.abs.thresh.1}, \code{days.below.1st}).}
\label{tab:hwcharacteristics}
\end{table}

The user can extensively customize the definition used to identify extreme events. The default is a definition for heat waves (two or more days with temperatures at or above the city's $98^{th}$ percentile of year-round temperature), but users can write a custom R function with either a different heat wave definition or with a definition appropriate for a different type of extreme event (e.g., one or more days at or below $0^oC$ for frost day spells). 
Some of these heat wave characteristics are based on relative temperature, which is a measure of how the temperature during a heat wave compares to the typical distribution of temperature at that location \citep{IPCCch1}. These relative temperature metrics will vary depending on whether you calculate them based on a location's present-day temperature distribution or on the location's temperature distribution in the future, since typical temperature distributions in many locations are expected to change with climate change. The package therefore allows the user to specify date ranges of the temperature distributions to be used in calculating these relative temperature metrics in each location.

Second, once the user creates these data frames of location-specific heat waves, the `apply_all_models` function allows the user to apply a custom function across all the heat wave data frames that were generated to create a summary data frame (Figure \ref{fig:overview}, right). This functionality can be used to generate summary statistics (e.g., determine average heat wave length or total heat wave days), or can be used to apply more complex functions (e.g., apply epidemiologic effect estimates across the heat waves to generate health impact estimates).

## Example data

We have included data files in the package to serve as example files, so users can try this package before applying it to their own directory of climate projection files. The example files are all included as comma-separated (.csv) files, rather than as saved R objects, because the `gen_hw_set` function requires a directory of comma-separated files as input and csv is a general file format familiar to most users. This example data therefore also allow users to see how to create the appropriate directory set-up for this package.

This example data comes from two climate models that are a part of CMIP5: (1) the model of the Beijing Climate Center, China Meteorological Administration (BCC) \citep{xin2013introduction} and (2) the National Center for Atmospheric Research's (NCAR's) Community Climate System Model, version 4 (CCSM4) \citep{gent2011community}. We include one ensemble member from BCC (r1i1p1) and two from CCSM (r1i1p1 and r2i1p1). To ensure that the size of this example data is reasonably small, we have only included projection data for grid points from these climate models that are near five U.S. east coast cities: New York, NY; Philadelphia, PA; Newark, NJ; Baltimore, MD, and Providence, RI. Further, to keep the file sizes reasonably small, the historical projections range over the years 1990 to 1999, while the future projections are limited to 2060 to 2079. Users' applications of this package will likely use directories with many more climate model ensemble members and more locations; however, the operation of the package is the same for this smaller example application as it would be for a much larger application.

Once the \pkg{futureheatwaves} package is installed and loaded, the user can find the local location of these files using R's `system.file` function. In the later sections of this article, we show how to use the package functions with these example files as inputs.

When setting up to run climate model output data beyond this example data, any downloaded CMIP5 files will require some pre-processing, in terms of getting the getting the data into a specific format and storing these files in a specific directory structure. A file of study locations must also be saved locally in a specific format. Extensive details on the required format of files and required input directory structure is available in the \pkg{futureheatwaves} vignette "futureheatwaves", while tips on processing CMIP5 files into the required structure are given in the package's "starting_from_netcdf" vignette.

## Basic example of processing projection files

Once these files and directories are set up, \pkg{futureheatwaves} can process them to identify and characterize heat waves in each ensemble member's projection for each  location using the `gen_hw_set` function. For example, to process the example climate projections directory included with the package, the user can run:

```{r eval = FALSE}
projection_dir_location <- system.file("extdata/cmip5",
                                       package = "futureheatwaves")
city_file_location <- system.file("extdata/cities.csv",
                                  package = "futureheatwaves")

gen_hw_set(out = "example_results",
           dataFolder = projection_dir_location ,
           dataDirectories = list("historical" = c(1990, 1999),
                                        "rcp85" = c(2060, 2079)),
           citycsv = city_file_location,
           coordinateFilenames = "latitude_longitude_NorthAmerica_12mo.csv",
           tasFilenames = "tas_NorthAmerica_12mo.csv",
           timeFilenames = "time_NorthAmerica_12mo.csv")
```

This code first identifies and saves as objects the path names on the user's computer of the example climate projections directory (`projection_dir_location`) and the file of study locations (`city_file_location`). The `gen_hw_set` function processes this example input and creates a new directory, `example_results`, with files of identified and characterized heat waves, in the user's current working directory. In this example code, this processing is done using  default values for the heat wave definition, years for which to generate the heat wave data sets, etc. Ways to customize these choices are fully explained later in the text.

In this call, the user must specify the directory where the results should be written (`out`), the location of the directory of climate projections (`dataFolder`), the names of the two main subdirectories of that climate projection directory, as well as their year boundaries (`dataDirectories`; see, for example, Figure \ref{fig:directories}), the location of the city location file (`citycsv`), and the names used for the grid coordinate, climate projection, and projection date files (`coordinateFilenames`, `tasFilenames`, and `timeFilenames`). When `gen_hw_set` is run, the user is advised that the function will write files to his computer and must agree to proceed:

```
 Warning: This function will write new files to your computer in the 
 ~/tmp/example_results/ directory of your computer. If that directory already exists,
running this function will write over it. 
 Do you want to continue? (y / n): 
```

\noindent The function provides status reports while processing. This helps the user see that the function call is progressing, as `gen_hw_set` can take a while to run if it is processing many locations and / or many climate ensemble members. Once the function has completed running, results will be written locally to the directory specified by the `out` argument of `gen_hw_set`. The structure of this output directory is shown in Figure \ref{fig:directories} (right). This directory will include files with some basic information about the climate models and the closest grid points of each climate model to each location. The output directory will also include a directory with files of identified and classified heat waves for each ensemble member, including all characteristics in Table \ref{tab:hwcharacteristics}.

## Customizing the extreme event definition

The default heat wave definition for this package is:

> A \dfn{heat wave} is two or more days at or above a city-specific threshold temperature, with the threshold determined as the $98^{th}$ percentile of year-round temperature in the city during some reference period (by default, 1990--1999).

Many different definitions of heat waves exist (e.g., \citet{smith2013heat, kent2014heat}), so researchers will often want to use alternative ways to define heat waves. Researchers might want to use a specific definition, for example, because it matches the definition used by local health officials to declare heat wave warnings or, in the case of health impact assessments, to match with a definition used in an epidemiological study. Three components of the heat wave definition can be easily customized in the `gen_hw_set` function call, without creating a new R function to use to identify heat waves. The customization of the heat wave definition is even more extensive as one has the option of writing a custom R function.

First, the percentile used to identify a heat wave can be changed using the `probThreshold` option in `gen_hw_set`. This option can take values between 0 and 1. The default value is 0.98, or a definition with a threshold of the $98^{th}$ percentile of the location's temperature. For example, to identify heat waves as two or more days at or above the city's $99^{th}$ percentile of year-round temperature, the user could run:

```{r eval = FALSE}
gen_hw_set(out = "example_results",
           dataFolder = projection_dir_location ,
           dataDirectories = list("historical" = c(1990, 1999),
                                        "rcp85" = c(2060, 2079)),
           citycsv = city_file_location,
           coordinateFilenames = "latitude_longitude_NorthAmerica_12mo.csv",
           tasFilenames = "tas_NorthAmerica_12mo.csv",
           timeFilenames = "time_NorthAmerica_12mo.csv",
           probThreshold = 0.99)
```

\noindent Second, the user can change the number of days used in the heat wave definition using the `numDays` argument in the `gen_hw_set` function. Combined, these two customization choices can allow the user to identify heat waves using many of the heat wave definitions used in previous climate and health research-- for example, 9 of the 16 heat wave definitions used in \citet{kent2014heat} could be fit using different combinations of these two options for specifying threshold percentile and number of days.

Third, it is possible to specify the range of years that should be used when determining this threshold. For example, if you wanted to base the threshold for each location on its current climate, you could leave this option as its default, while if you wanted to use a different set of years (i.e. a threshold relative to future projected temperatures), you could set the start and end year bounds for this reference period using the `thresholdBoundaries` argument in the function `gen_hw_set`. For example, to use temperatures in each city from 2070 to 2079 to determine threshold temperatures, you would run:

```{r eval = FALSE}
gen_hw_set(out = "example_results",
           dataFolder = projection_dir_location ,
           dataDirectories = list("historical" = c(1990, 1999),
                                        "rcp85" = c(2060, 2079)),
           citycsv = city_file_location,
           coordinateFilenames = "latitude_longitude_NorthAmerica_12mo.csv",
           tasFilenames = "tas_NorthAmerica_12mo.csv",
           timeFilenames = "time_NorthAmerica_12mo.csv",
           thresholdBoundaries = c(2070, 2079))
```

It is also possible to use a completely customized heat wave definition. This functionality allows the user to use heat wave definitions that either require a number of days above an absolute threshold (e.g., maximum temperature of $\ge95^oF$ for $\ge1$ day; \citet{kent2014heat, tan2007heat}) or that require a combination of thresholds to be met (e.g., maximum daily temperature above a lower threshold every day of the heat wave and above a higher threshold for a certain number of days; \citet{kent2014heat, peng2011toward}).

To use a customized heat wave definition, the user can write and load an R function that implements the definition. To work correctly, this custom function must allow only specific inputs and generate only specific outputs. Details about this required structure are provided in the \pkg{futureheatwaves} package vignette.

To help in developing custom heat wave identification functions, the package includes an example of the required input data frame, `datafr`, which can be loaded with `data(datafr)`. If a custom heat wave function is properly set up to use in this package, it can take as input this example data frame and a threshold value and will return the original data frame with the columns for whether the day was part of a heat wave (`hw`) and, if so, the sequential number of the heat wave out of all heat waves in that location (`hw.number`).

The custom function can then be referenced using the `IDheatwavesFunction` option in `gen_hw_set`, and it will be used to identify heat waves in the data set. As an example, we have included an alternative ID definition function in this package, called `IDHeatwavesAlternative`, which identifies heat waves as two or more days above the higher of either the location's $98^{th}$ percentile temperature or $80^{o}F$, with the idea that this definition prevents the identification of heat waves at mild temperatures in locations with mild or cool climates. The following code processes the example climate projecting files using this function as the heat wave definition:

```{r, eval = FALSE}
gen_hw_set(out = "example_results",
           dataFolder = projection_dir_location ,
           dataDirectories = list("historical" = c(1990, 1999),
                                        "rcp85" = c(2060, 2079)),
           citycsv = city_file_location,
           coordinateFilenames = "latitude_longitude_NorthAmerica_12mo.csv",
           tasFilenames = "tas_NorthAmerica_12mo.csv",
           timeFilenames = "time_NorthAmerica_12mo.csv",
           IDheatwavesFunction = "IDHeatwavesAlternative")
```

With this ability to write and use a custom heat wave identification function, most heat wave definitions can be applied with this package. The exception is definitions that require two separate temperature metrics (e.g., maximum temperature above a certain threshold or minimum temperature above a different threshold; \citet{kent2014heat, robinson2001definition}), because the `gen_hw_set` function cannot take as input projections of more than one temperature metric. Definitions that use a heat metric that combines multiple inputs (e.g., heat index, which is a combined measure of temperature and air moisture) can be used, but require that the calculation of the metric is done when creating the input files to include in the ``Climate projections" directory (Figure \ref{fig:overview}, left), rather than as part of the heat wave definition function.

To increase processing speed when identifying heat waves, we coded parts of the default heat wave identification function in C++ and synced it with R using the \CRANpkg{Rcpp} package \citep{Rcpp}. Users should consider a similar strategy for custom heat wave definitions, especially if processing a large number of climate projection files.

## Exploring extreme events

Once you have created a directory of files with characterized heat waves for each ensemble member, the results can be explored using the `apply_all_models` function. This function allows the user to apply custom R functions across all heat wave data frames created by the `gen_hw_sets` call. The user can apply any R function that follows certain standards in accepting input and returning output. Full details on these standards are given in the \pkg{futureheatwaves} package vignette.

As an example, if the user wanted to to get the average temperature of the heat waves identified within each ensemble member, he or she could write a simple function:

\begin{verbatim}
average_mean_temp <- function(hw_datafr){
        out <- mean(hw_datafr$mean.var)
        return(out)
        }
\end{verbatim}

\noindent The `apply_all_models` function can then apply this `average_mean_temp` function across the heat wave data frames for all ensemble members in all climate models:

```{r echo = FALSE}
average_mean_temp <- function(hw_datafr){
        out <- mean(hw_datafr$mean.var)
        return(out)
        }
```

```{r}
out <- system.file("extdata/example_results", package = "futureheatwaves")
apply_all_models(out = out, FUN = average_mean_temp)
```

Note that the location of the directory with the heat wave data frames must be specified using the `out` argument when calling \code{apply\_all\_models}. Typically, this will be the directory path for the directory specified with the `out` argument in \code{gen\_hw\_set}. Location-specific results can be generated using the \code{city\_specific} argument in \code{apply\_all\_models}:

```{r}
apply_all_models(out = out, FUN = average_mean_temp, city_specific = TRUE)
```

This output is structured as ``tidy" data \citep{wickham2014tidy}, allowing it to be used easily with the graphing package \CRANpkg{ggplot2} \citep{ggplot2}.

The `apply_all_models` function can also be used to project the health impacts of heat waves. As a very simplistic example, \citet{anderson2009weather} estimated that heat waves, defined as two or more days at or above a community's $98^{th}$ percentile temperature, were associated with an added relative risk of 1.032 for cardiorespiratory mortality risk in 107 U.S. communities. A simple estimate of excess deaths associated with this added heat wave risk in a community can be calculated as \citep{peng2011toward}:

\begin{equation}
E_c = N_c * (RR - 1) * L_c 
\label{eq:healthimpacts}
\end{equation}

where:

- $E_c$ is the total number of excess deaths in community $c$;
- $N_c$ is the baseline average daily mortality in community $c$; 
- $RR$ is the relative risk of cardiorespiratory mortality per day associated with a heat wave; and 
- $L_c$ is the total number of heat wave days in community $c$ over the study period.

This impact assessment equation can be translated into a function that merges each projection's heat wave data frame with a data frame of community-specific baseline mortality rates ($B_c$), calculates equation \ref{eq:healthimpacts} for each heat wave, and then sums up the total excess deaths across all heat waves:

\begin{verbatim}
excess_deaths <- function(hw_datafr, base_mortality, RR = 1.032){
        hw_datafr <- dplyr::left_join(hw_datafr, base_mortality,
                                      by = "city") %>%
                dplyr::mutate(excess_deaths = base_mort * length * RR)
        out <- sum(hw_datafr$excess_deaths)
        return(out)
}
\end{verbatim}

\noindent Once defined in R, this function can be applied across all heat waves from all climate models' ensemble members, provided that you have a data frame called `city_mortality` with columns with each community's identifier (`city`) and baseline mortality rate (`base_mort`), using the call:

```{r eval = FALSE}
apply_all_models(out = out, FUN = excess_deaths, base_mortality = city_mortality)
```

As examples of the type of functions that can be used with `apply_all_models`, we have included several simple functions for exploring heat waves (Table \ref{tab:applyfunctions}). We included an example data frame representative of the heat wave data frames that `gen_hw_set` outputs, to use to test new functions. This data can be loaded with `data(hw_datafr)`. The process for creating a new function to use to explore heatwaves should be to:

1. Load the example data frame with `load(hw_datafr)`;
2. Build a function that works with `hw_datafr` as an input; and 
3. Apply the function with `apply_all_models` to process all of the heat wave files.

\begin{table}
\centering
\begin{tabular}{ll}
\toprule
Function & Description \\
\midrule
\code{number\_of\_heatwaves} & Determines the number of extreme events\\
\code{heatwave\_days} & Sums up the total number of extreme event days across all heat waves\\
\code{average\_length} & Calculates the average length of extreme events\\
\code{average\_mean\_temp} & Calculates the average mean temperature across all extreme events\\
\bottomrule
\end{tabular}
\caption{Example functions included in the \pkg{futureheatwaves} package that can be used to explore extreme events using the \code{apply\_all\_models} function.}
\label{tab:applyfunctions}
\end{table}

## Customizing dates for projections and reference temperatures

By default, the `gen_hw_set` function will generate heat wave projections for the years 2070 to 2079, to align with the example data. This projection date range can be changed by specifying alternative starting and ending years in the `projectionBoundaries` argument of the `gen_hw_set` function. For example, to create projections for 2060 to 2079, the user would call:

```{r, eval = FALSE}
gen_hw_set(out = "example_results",
           dataFolder = projection_dir_location ,
           dataDirectories = list("historical" = c(1990, 1999),
                                        "rcp85" = c(2060, 2079)),
           citycsv = city_file_location,
           coordinateFilenames = "latitude_longitude_NorthAmerica_12mo.csv",
           tasFilenames = "tas_NorthAmerica_12mo.csv",
           timeFilenames = "time_NorthAmerica_12mo.csv",
           projectionBoundaries = c(2060, 2079))
```

The heat wave data sets characterize heat wave in several ways that are based on relative temperature (for example, the `mean.temp.quantile`, `max.temp.quantile`, and `min.temp.quantile` columns described in Table \ref{tab:hwcharacteristics}). These characteristics are measured for each of the heat waves identified by the `gen_hw_set` function by taking the absolute temperature of the heat wave (e.g., average temperature during the heat wave is $90^oF$) and comparing it to the heat wave location's typical temperature distribution. This process generates relative measures of how intense the heat wave is compared to what is normal in that location (e.g., $90^oF$ is in the $99^{th}$ percentile of year-round temperatures in the location).

The values of these relative heat wave characteristics will depend on what time period is used to determine the location's typical temperature distribution. For example, the same heat wave would likely have much higher values for these relative characteristics if the temperature during the heat wave is compared to typical temperatures in the present day for that location versus if the temperature is compared to typical temperatures much later in the $21^{st}$ century for the same location.

The default in the `gen_hw_set` function is to use temperatures for the period 2070 to 2079 for these reference temperatures to calculate all relative heat wave characteristics. However, the user can change this specification using the `referenceBoundaries` option of `gen_hw_set`. This functionality can be useful in exploring the role of adaptation in future heat waves. For example, to use temperature projections for 1990 to 1999 when calculating relative characteristics of heat waves, to explore the assumption that cities remain adapted to their present-day climate, rather than changing in adaptation as climate change increases temperatures, a user could run:

```{r, eval = FALSE}
gen_hw_set(out = "example_results",
           dataFolder = projection_dir_location ,
           dataDirectories = list("historical" = c(1990, 1999),
                                        "rcp85" = c(2060, 2079)),
           citycsv = city_file_location,
           coordinateFilenames = "latitude_longitude_NorthAmerica_12mo.csv",
           tasFilenames = "tas_NorthAmerica_12mo.csv",
           timeFilenames = "time_NorthAmerica_12mo.csv",
           referenceBoundaries = c(1990, 1999))
```

For these date range specifications, there are some restrictions on which year ranges can be selected. The starting year cannot be earlier than the first year in the first subdirectory, the ending year cannot be later than the last year of the second subdirectory, and the custom date boundaries cannot span the two subdirectories.

## Mapping grid points

It can be useful to explore the location of the climate model grid point used to pull climate model output for each study location with a given climate model. Therefore, the package has a function called `map_grid_leaflet` that plots the locations of grid points used for each location from each climate model. This function is built using the htmlWidget \pkg{leaflet} package \citep{leaflet}. The following code illustrates the use of this function with the example data to create Figure \ref{fig:gridmap}, which plots the grid points used in the example data from the BCC climate model in the example data:

```{r eval = FALSE}
out <- system.file("extdata/example_results", package = "futureheatwaves")
map_grid_leaflet(plot_model = "bcc1", out = out)
```

\begin{figure}
\begin{center}
\includegraphics[width = 0.9\textwidth]{ExampleLeaflet}
\end{center}
\caption{Snapshot of an example interactive map created using the \code{map\_grid\_leaflet} showing the locations of study cities and their matching climate model grid points for the BCC climate model example data included with \pkg{futureheatwaves}. The lines on the map connect each climate model grid point to the study location(s) for which that grid point was used. The interactive maps include pop-ups with city identiers; one is shown open in this snapshot as an example. From this map, you can see that the climate model grid point closest to New York City for this climate model is over the Atlantic Ocean.}
\label{fig:gridmap}
\end{figure}

## Extensions

The functionality of this package can be easily expanded by loops. For example, to explore the role of the heat wave definition on projections, the user could create a loop to run `gen_hw_set` and `apply_all_models` to the same directory of climate projections but with a variety of different functions used to identify the heat waves in the projections.

Also, while this package was created to be used for research on heat waves in climate change projections, with some modifications it can be used more broadly. For example, there are other episodes like wildfires and air pollution where it may be interesting to identify extended periods of high exposures in projection time series, and this package could be applied to explore these exposures. The directory of projection data would need to be set up in the same structure as for exploring heat waves, and the `input_metric` should be set as `fahrenheit`, to pass the exposure values through to the characterized data sets without performing a conversion. A user could also use this package to explore events that must be lower than some minimum threshold (e.g., cold waves), but it would take some extra coding and conversions, since the functions in this package are written to identify periods above a threshold (for example, the user could multiple all projected temperatures by -1, and then the coldest temperatures would register as being the highest).

```{r eval = FALSE}
## Cold wave example
gen_hw_set(out = "example_results",
           dataFolder = projection_dir_location ,
           dataDirectories = list("historical" = c(1990, 1999),
                                  "rcp85" = c(2060, 2079)),
           citycsv = city_file_location,
           coordinateFilenames = "latitude_longitude_NorthAmerica_12mo.csv",
           tasFilenames = "tas_NorthAmerica_12mo.csv",
           timeFilenames = "time_NorthAmerica_12mo.csv", 
           probThreshold = 0.10, 
           above_threshold = FALSE, 
           absolute_thresholds = c(266, 263, 260, 258), 
           seasonal_months = c(12, 1, 2))
```

>  Extreme statistics and spell-lengths: "A 'frost day' is defined as one during which the minimum temperature falls below freezing point (0 degC). This is described as a climatological statistic, in which the minimum temperature is first calculated within each day, and then the number of days or spell lengths meeting the specified condition are evaluated." (from one example in http://cfconventions.org/Data/cf-conventions/cf-conventions-1.6/build/cf-conventions.html#calendar)

Users doing these kinds of extensions will need to pay attention to a few points. First, some of the event characteristics (first in the calendar year, average of May--September temperatures, days above $90^{o}F$) might not be meaningful for studies of other types of events. Further, because event periods are usually defined as a string of multiple days exceeding some threshold, the functions in this package may miss the first and last event of the time period. For example, if the first day of the time series is the last day of an event, this function would not identify that event because it lacks data from the earlier days that allow this day to meet the event definition. This issue would lead to, at most, missing two events out of each projection, but should be considered if studying events that might occur near the start and end of projection data. If there is adequate interest from researchers, in the future we may adapt the package to make these secondary applications part of the package.

# Acknowledgements

This work was supported by grants from the National Institute of Environmental Health Sciences (R00ES022631), the National Science Foundation (1331399), and the Colorado State University Vice President for Research.

\bibliography{Anderson}