### 2016-12-30 paper_drafts/RJournalRevisions/anderson-eason-barnes.tex

% !TeX root = RJwrapper.tex
\title{Working with daily climate model output data in R and the
\pkg{futureheatwaves} package}
\author{by G. Brooke Anderson, Colin Eason, Elizabeth A. Barnes}

\maketitle

\abstract{%
Research on climate change impacts can require extensive processing to
incorporate output from ensembles of different climate models and
different simulations of each model. This processing can be particularly
extensive when identifying and characterizing extreme events like heat
waves and frost day spells, as these must be processed from model output
with daily time steps and may use identification threholds based on
local distributions (e.g., 98\textsuperscript{th} percentile temperature
for that location). Here, we provide an overview of working with daily
climate model output data in R. Further, we present the
\pkg{futureheatwaves} package, which we developed to ease the process of
identifying, characterizing, and exploring multi-day extreme events in
climate model output. This package takes as input a directory of
projection files from one or many climate models, identifies all
occurrences of a specific type of extreme event (e.g., heat waves) for
multiple cities using a user-specified event definition, and writes an
output directory of files with these extreme events and their
characteristics (e.g., length, absolute and relative temperatures). The
\pkg{futureheatwaves} package then allows the user to create custom
functions and apply them across all event files; this functionality can
be used either to explore extreme event patterns (e.g., average length
and intensity across different climate models) or to apply
epidemiological estimates to project health impacts under different
scenarios.
}

\section{Introduction}\label{introduction}

Research on climate change impacts can require extensive processing.
This is not only because output files for a single climate model can be
large, but also because of the rising popularity of ensemble techniques
\citep{IPCCch1}, in which, to better characterize uncertainty in
projections, impacts are assessed for multiple climate models, multiple
simulations of each climate model, and multiple climate experiments.
Projections of regional climate change over the next century are subject
to high uncertainty due to three distinct sources: (1) internal climate
variability, i.e.~climate noise, (2) climate model uncertainty, i.e.~the
same forcing can produce a different response in different models and
(3) scenario uncertainty, i.e.~uncertainty in future climate forcings
(e.g. \citet{hawkins2009potential}). One approach for ensuring that
these sources of uncertainty are characterized is to simulate the future
climate many times with multiple models and for multiple future
scenarios.

For example, the Coupled Model Intercomparison Project, phase 5 (CMIP5;
\citet{taylor2012overview}) brought together dozens of major climate
modeling groups around the world to simulate the same future radiative
forcing scenarios, but with their own models. This created an ensemble
of state-of-the-art climate model projections that allow researchers to
study projections and their uncertainties. Most of these modeling groups
additionally performed more than one simulation for each scenario and
model (i.e.~multiple ensemble members), perturbing the initial
conditions by a very tiny amount to quantify uncertainties due to
internal climate variability.

Processing climate output data is particularly intensive for data in
smaller time steps, like daily output. While some climate impacts can be
assessed using climate model output at monthly, seasonal, or yearly time
steps, the impacts of multi-day extreme events must be assessed using
output in daily time steps. Such multi-day extreme events include heat
waves, cold spells, frost day spells, and droughts. The assessment is
further complicated if extreme events are identified based on conditions
that are rare for a certain location (e.g., 98th percentile of local
temperature distribution for identifying heat waves) \citep{IPCCch1}, as
this event definition must be determined at each study location from
climate model output. Further, it is often of interest to create
summaries of multiple characteristics of these extreme events. For
example, one may be interested in determining whether the frequency or
characteristics (e.g., length, intensity) of heat waves or warm spells
will change under certain climate change scenarios \citep{IPCCch1}.

We begin this article with an overview of climate model output data,
particularly daily data, for R users. We outline where data from CMIP5
can be obtained as well as how to work with the file format (netCDF)
from R. We overview some R packages that can be useful when working with
this data, as well as aspects of the data (e.g., non-standard calendars)
of which users should be aware when working with daily climate model
output in R.

After this overview, we present the \pkg{futureheatwaves} package which
we created to aid in identifying and characterizing any type of
multi-day extreme event from daily climate model output (Table
\ref{tab:goals}). This package can be used to identify and characterize
a variety of multi-day extreme events, but provides some functionality
particularly useful in identifying and characterizing heat waves
specifically. Quantification of the impacts of heat waves on human
health suffer from additional sources of uncertainty beyond those
inherent in projections of regional changes in surface temperature.
These include: (1) uncertainty in the ability of communities to adapt to
changing temperatures (the adaptation scenario) and (2) uncertainty in
the definition of a heat wave itself. Thus, identifying and
characterizing the impacts of future heat waves in state-of-the-art
climate models requires analyzing hundreds of projections based on
combinations of anthropogenic activity, climate model and ensemble
member, adaptation scenario, and heat wave definition. Such an analysis
is non-trivial. For example, to estimate the possible impacts of heat
waves using 32 ensemble members of CMIP5, two forcing scenarios, two
different heat wave definitions, and five progressive assumptions about
adaptation to heat, would require that one identify and characterize
heat waves in 640 separate temperature time series for each region of
interest.

\begin{table}[t]
\begin{center}
\newcolumntype{L}[1]{>{\raggedright\arraybackslash}p{#1}}
\begin{tabular}{lL{\dimexpr0.8\textwidth-2\tabcolsep\relax}}
\toprule
 & Design goals of the \pkg{futureheatwaves} package \\
\midrule
1 & Make processing of large sets of climate projections more practical for researchers exploring the potential impacts of heat waves and other multi-day extreme events.\\
2 & Speed up processing time by incorporating C++ in event identification.\\
3 & Keep track of the names of climate models, and number of ensemble members processed for each.\\
4 & Not only identify, but also characterize, all extreme events within each climate projection, to allow the exploration of patterns in these characteristics across different projections and also to allow the use of more complex impacts models, including models that include effect modification by event characteristics (e.g., event length, event intensity). For example, this package allows the user to apply a health effects model where risk of mortality is not the same for every heat wave, but rather is modified by heat wave length, intensity, or other measured characteristics.\\
5 & Give users extensive power in customizing the process, including allowing custom event definitions.\\
6 & Allow users to easily explore the extreme events identified within all climate projections by applying custom functions across heat wave data sets from all projections at once.\\
7 & Create output that is in a ``tidy" data format, allowing it to work well with \pkg{ggplot2} for visualization.\\
\bottomrule
\end{tabular}
\end{center}
\caption{Design goals for the \pkg{futureheatwaves} package.}
\label{tab:goals}
\end{table}

The \pkg{futureheatwaves} package automates the process of identifying
and characterizing multi-day extreme events across different ensemble
members of one or more climate models. A variety of different heat wave
definitions have been used to identify heat waves in a time series of
temperature data \citep{smith2013heat}, and the choice of heat wave
definitions can influence both health effect estimates
\citep{chen2015influence, kent2014heat} and projected heat wave trends
\citep{smith2013heat}. Further, other types of extreme events will be
defined differently than heat waves (for example, frost day spells may
be defined as one or more days with temperature at or below
32\textsuperscript{o}F / 0\textsuperscript{o}C). This package therefore
allows the user to create and use a custom extreme event definition to
identify events in the climate model output. Finally, this package
allows the user to explore the extreme events identified in the climate
model output with a function that can take a user-defined R function and
apply it across all extreme event files generated for the separate
ensemble members.

\section{An overview of climate model output for R
users}\label{an-overview-of-climate-model-output-for-r-users}

\subsection{CMIP5 climate model output
data}\label{cmip5-climate-model-output-data}

For climate impact studies, a main source for climate model output is
the Coupled Model Intercomparison Project, which is currently in its
fifth phase (CMIP5). Over 20 climate modeling groups have created one or
more climate models which, for this project, are run using standardized
scenarios \citep{taylor2012overview}. The resulting output is uniform
across modeling groups and has a consistent structure, which allows
comparison of simulations from different models \citep{IPCCch9}. For
CMIP5, each group ran simulations under several experiments, with
experiments varying in terms of the radiative forcing. This radiative
forcing depends on time-varying model inputs (greenhouse gas emissions
or concentrations, land use changes, etc.), which are specified for each
experiment \citep{taylor2012overview, IPCCch9}. Experiments include
historical experiments (run using forcings consistent with observed and
reconstructed data for 1850--2005), pre-Industrial control experiments,
and experiments of future scenarios of radiative forcing over the 21st
century or longer (e.g., RCP4.5, RCP8.5) \citep{taylor2012overview}.
Some modeling groups created ensembles of output for a specific model
and experiment, in which they ran the experiment multiple times with the
model with very small changes to the initial conditions, resulting in
multiple ensemble members for a single climate model and experiment.
CMIP5 climate model output is archived at a number of different time
steps (e.g., daily, monthly, seasonal, yearly) \citep{taylor2010cmip5},
and some variables are reported at multiple levels in the ocean or
atmosphere (e.g., ocean temperature). Here, we will focus on data with a
daily time step for variables reported at a single level (e.g.,
near-surface air temperature).

The CMIP5 climate model output data is distributed across data nodes at
different climate modeling centers \citep{taylor2012overview}, but can
be accessed centrally at the World Climate Research Programme CMIP5 data
portal at \url{https://pcmdi.llnl.gov/search/cmip5/}. Users must
register before downloading data, and some data are restricted to
non-commercial use. There is a separate file for each combination of
climate model, experiment, modeling realm (e.g., atmosphere, ocean),
variable, time step, and ensemble member
\citep{taylor2012overview, taylor2010cmip5}. For finer time scales, this
output is further split across multiple files for specific year ranges
(e.g., 5 years of output for each file) \citep{taylor2010cmip5}.
Filenames for CMIP5 files can be parsed to generate information about
the output variable, climate model, experiment, and ensemble member for
the simulation \citep{taylor2010cmip5}. Files can be searched and
downloaded through a point-and-click web interface. The data portal also
allows you to generate a specific wget script, which you can use to
download many files at once. A wget script can also be generated through
Earth System Grid Federation's Search RESTful API. Tips on efficiently
searching and downloading the data, including through use of wget
scripts and the search API, are available as user tutorials through the
website of the University of Colorado Boulder's Earth System CoG (e.g.,
\url{https://www.earthsystemcog.org/projects/cog/doc/wget} for a
tutorial on downloading files using wget).

CMIP5 files are saved in Network Common Data Format (netCDF). NetCDF is
a binary file format that allows storage of data representing a regular
array. These climate model output files can be as large as several
gigabytes {[}?{]}. For climate model output at a single level (e.g.,
near-surface air temperature), the data is a 3-dimensional array, with
dimensions representing time and two coordinates of location (e.g.,
latitude and longitude) (Figure \ref{fig:netcdfexample}). Global climate
models generate output at regularly-spaced time steps, typically at
regularly-spaced grid points around the world. The latitude and
longitude spacing of grid points vary by climate model, but are
typically 1--2 degrees for atmospheric variables in CMIP5 models
\citep{IPCCch9}. Each data point in the netCDF array gives the modeled
value of the variable (e.g., surface temperature) for a single time
point and location. For CMIP5 climate model output, the location units
are in degrees east and degrees north for longitude and latitude,
respectively. For daily output files, the time unit is in days since a
specified origin date-time (e.g., days since 1850-01-01 00:00:00)
\citep{taylor2010cmip5}. All CMIP5 output files are required to include
certain metadata \citep{taylor2010cmip5}. This required metadata
includes the experiment, forcing agents input to the model to create the
simulation, time step, institution and institutional contact
information, climate model, and modeling realm \citep{taylor2010cmip5}.
The metadata also must include units for all of the coordinate variables
(e.g., longitude, latitude, time).

\begin{widefigure}
\begin{center}
\includegraphics[width = \textwidth]{netcdf_figure}
\end{center}
\caption{Example of structure of a NetCDF climate model output file for a variable reported at a single depth, like surface air temperature. Data are stored in a three-dimensional array, with measurements at each time step and grid location. Surface temperature data are typically indexed in climate model output files by longitude, latitude, and time, in that order. For example, if the air surface temperature ("tas") is read into an R object called \code{tas}, you can access the value for the first day at the fourth longitude and third latitude with \code{tas[4, 3, 1]}. In addition to the output variable (temperature in this example), vectors with the ordered values of each dimension (longitude, latitude, and time) can also be read in from the netCDF file, as well as attribute data (e.g., units for variables, the calendar used for time).}
\label{fig:netcdfexample}
\end{widefigure}

\citet{taylor2012overview} and \citet{meehl2007wcrp} are excellent
resources for finding out more about the CMIP climate model output data.

\subsection{Working with climate model output in
R}\label{working-with-climate-model-output-in-r}

A few R packages can be used to work with the netCDF file format used
for CMIP5 files. Earlier packages to work with netCDF files included
\pkg{ncdf} and \pkg{ncvar}, but these do not work with the newer netCDF
version 4 released in 2008 and are no longer available through CRAN.
More recent packages, including \pkg{ncdf4} \citep{ncdf4} and
\pkg{RNetCDF} \citep{michna2013rnetcdf, RNetCDF}, work with both version
4 and netCDF's older version 3. Climate model output data for CMIP5 is
required to conform with the earlier version (version 3)
\citep{taylor2010cmip5} and so should work with any of these packages,
although it is safer to write code using functions that can be used with
either version in case future phases of CMIP do not require files to
conform with netCDF version 3.

The netCDF format allows you to access metadata and variables describing
the dimensions of the data without reading the full file into memory.
The metadata describes the dimensions of the array and the units, etc.,
of each variable and is printed by the print method of the R object
returned by the \code{nc\_open} function of \pkg{ncdf4}. For CMIP5
files, the dimension variables include the times and locations
corresponding to each array in the data. Before reading in a variable
from a netCDF file, a connection to the file must be opened, for example
with the \code{nc\_open} function from \pkg{ncdf4}. Variables can then
be read in using the \code{ncvar\_get} function from \pkg{ncdf4}, with
the \code{varid} parameter set to ``lat'', ``lon'', or ``time'' (as a
caveat, many climate models output to non-Gregorian calendars, in which
case the time variable should be read in using a different function, as
discussed later in this section). The climate output variable (e.g.,
near-surface air temperature) can similarly be read in using
\code{ncvar\_get}. In this case, the \code{varid} parameter should be
set using the appropriate CMIP5 variable name (e.g., ``tas'' for
near-surface air temperature); these variable names can be found in the
CMIP Requested Output tables \citep{taylor2010cmip5}. You can use the
dimensional time and location data to identify the location of the
variable data you need in the netCDF array and use indexing to read only
that data into memory, without needing to read in the full file (for
example, with the \code{nc.get.var.subset.by.axes} function in
\pkg{ncdf4.helpers}). Once the user is done reading in data from the
file, the connection can be closed (e.g., with the \code{nc\_close}
function from \pkg{ncdf4}).

Since the late 1500s, Western dates have been set using the Gregorian
calendar, which has 365.2425-day years. Some climate models, however,
are run using different calendars, including the Julian calendar
(365.25-day years), a calendar where there are no leap years (365-day
years), a calendar where every year is a leap year (366-day years), and
a calendar of twelve 30-day months (360-day years)
\citep{cfconventions}. With these non-Gregorian calendars, R's base
functions for converting a vector to a Date class based on the number of
days since an origin date (\code{as.Date}, \code{as.POSIXct}) do not
return the desired values. The \pkg{PCICt} \citep{PCICt} and
\pkg{ncdf4.helpers} \citep{ncdf4.helpers} packages provide further
functionality with netCDF files that can be useful when working with
climate model output data and provide particular help in working with
different calendars. CMIP5 netCDF files include information on the
calendar used, which the \code{nc.get.time.series} function in
\pkg{ncdf4.helpers} use to convert the ``time'' variable in the file to
an object of the PCICt class, which provides Date-like functionality for
360- and 365-day calendars \citep{PCICt}. As a note, while these
functions will help with handling most CMIP5 files, the CMIP5 standards
allows use of other calendars which may not be succesfully handled by
these functions, so it is important to assess whether the time variable
range in the PCICt object correctly matches the expected date ranges for
a file as you process CMIP5 data in R.

The size of the climate model output files can be large enough that it
may make more sense to work with smaller chunks of the data, rather than
reading all data into memory and working with the data all at once
\citep{RCMIP5}. This problem aggregates when working with multiple
climate models and more than one ensemble member for each of those
climate models.

The functions and packages described can be used to work with CMIP5
netCDF files in R to do things like map near-surface air temperatures
from a single climate model on a specific day in July 2075 (Figure
\ref{fig:worldmap}) and pull a time series of daily near-surface air
temperature simulations for the climate model grid point closest to
Beijing, China (Figure \ref{fig:timeseries}). The
``starting\_from\_netcdf'' vignette that comes with the
``futureheatwaves'' package provides all code required to create these
figures, as well as more details and code examples on working with CMIP5
netCDF files in R.

\begin{figure}
\begin{center}
\includegraphics[width = \textwidth]{worldmapexample}
\end{center}
\caption{Example of mapping near-surface air temperature data worldwide for a single day of climate model output data. This map uses data from the Geophysical Fluid Dynamics Laboratory's Earth System Model 2G, r1i1p1 ensemble member, on a single day in the summer of 2075.}
\label{fig:worldmap}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[width = 0.8\textwidth]{timeseriesexample}
\end{center}
\caption{Example of plotting a time series of temperature simulations between 2071 and 2075 from CMIP5 daily climate model output data for the model grid cell point closest to Beijing, China. This plot uses data from the Geophysical Fluid Dynamics Laboratory's Earth System Model 2G, r1i1p1 ensemble member.}
\label{fig:timeseries}
\end{figure}

In addition to these general packages for working with netCDF files,
there are several R packages specifically for working with climate model
output data, including \pkg{RCMIP5} \citep{RCMIP5} and \pkg{wux}
\citep{wux}. However, these packages are more useful for working with
data output at time steps of a month or higher and have limited utility
with the daily climate model output data required for studies of
multi-day extreme events.

The \pkg{RCMIP5} package includes functions to read in CMIP5 data from
netCDF files, scan a directory of CMIP5 files and determine models with
continuous available data, create objects of a special ``cmip5data''
class to work with CMIP5 data within R, and parse the filenames for all
files in a directory to extract information within the filename. For
this package, most functions only work with monthly or less frequent
data (e.g., the functions \code{checkTimePeriod} and \code{cmip5data},
as well as functions that work with \code{cmip5data} objects including
\code{filterDimensions} and \code{getProjectionMatrix}) \citep{RCMIP5}.
While the \code{loadCMIP5} function does successfully load the daily
data as a \code{cmip5data} object, most of the methods for this object
type do not do anything meaningful for daily data. The package's
\code{getFileInfo} function, however, will work with CMIP5 files of any
time step; this function identifies all CMIP5 files in a directory and
creates a dataframe that parses the information contained in the file
name. As a note, the \code{get.split.filename.cmip5} function in the
\pkg{ncdf4.helper} package similarly can be used to parse information
contained in CMIP5 file names \citep{ncdf4.helpers}.

The \pkg{wux} package \citep{wux} includes functions that allow the user
to download CMIP5 monthly-aggregated output directly from within R with
the \code{CMIP5fromESGF} function. However, this function does not allow
downloading of climate model output with finer time steps, like daily
data. This package uses the \code{models2wux} function to read in
climate model output netCDF files and convert to a WUX dataframe that
can be used by other functions in the package. While this function can
input climate model output with daily time steps, if the element
``what.timesteps'' of the \code{modelinput} list input is set to
``daily'', the function aggregates this data to a monthly or less
frequent (e.g., seasonal) aggregation when creating the WUX dataframe.
Therefore, while this package provides very useful functionality for
working with averaged output of daily climate model output data, it
cannot easily be used to identify and characterize multi-day extreme
events like heat waves.

More tips on working with CMIP5 files in R are provided in the
``starting\_from\_netcdf'' vignette of the \pkg{futureheatwaves}
package.

\section{\texorpdfstring{The \pkg{futureheatwaves}
package}{The  package}}\label{the-package}

\subsection{How the package works}\label{how-the-package-works}

\begin{widefigure}
\includegraphics[width = 0.9\textwidth]{OverviewFigure}
\caption{Overview of the functionality of the \pkg{futureheatwaves} package. The package takes a directory with climate projection files (left), for one or more climate models, with one or more ensemble members for each climate model (this example figure shows four climate models with one or two ensemble members each). The gen\_hw\_set function processes these files to create a data frame for each ensemble member, identifying and characterizing all multi-day extreme events (e.g., heat waves) in the time series projection for that ensemble member. The apply\_all\_models function allows users to explore these extreme events by applying user-created functions across all the extreme event data frames, creating a summary data frame with results.}
\label{fig:overview}
\end{widefigure}

We created the \pkg{futureheatwaves} package to aid in identifying,
characterizing, and exploring multi-day extreme events in daily climate
model output data. Figure \ref{fig:overview} gives an overview of the
two primary functions of the \pkg{futureheatwaves} package. First, the
\code{gen\_hw\_set} function processes a directory of climate projection
files that are stored locally on the user's computer (Figure
\ref{fig:overview}, ``Climate projections''), to generate a list of all
extreme events in each projection, as well as over a dozen
characteristics of each identified extreme event (Table
\ref{tab:hwcharacteristics}). The extreme events are identified and
characterized at one or more study locations (e.g., cities), with study
locations specified by the user in an input file. The extreme events
identified for each ensemble member are output as separate files in a
directory specified by the user (Figure \ref{fig:overview}, ``Extreme
events datasets``).

Once the user creates these data frames of location-specific heat waves,
the \code{apply\_all\_models} function allows the user to apply custom
functions across all the extreme event data frames generated by
\code{gen\_hw\_set} to summaries of extreme events across all climate
models and ensemble members (Figure \ref{fig:overview}, right). This
functionality can be used to generate summary statistics (e.g.,
determine average heat wave length or total frost days), or can be used
to apply more complex functions (e.g., apply epidemiologic effect
estimates across the heat waves to generate health impact estimates).

\begin{table}
\newcolumntype{L}[1]{>{\raggedright\arraybackslash}p{#1}}
\begin{tabular}{lL{\dimexpr0.8\textwidth-2\tabcolsep\relax}}
\toprule
Column name & Description of characteristic \\
\midrule
\code{mean.var} & Average daily value of the variable across all days in the extreme event, in the units in which the variable is expressed in input files (e.g., average daily mean temperature during the heat wave in degrees Kelvin) \\
\code{max.var} & Highest daily value of the variable across all days in the extreme event, in the units in which the variable is expressed in input files \\
\code{min.var} & Lowest daily value of the variable across all days in the extreme event, in the units in which the variable is expressed in input files \\
\code{length} & Number of days in the event \\
\code{start.date} & Date of the first day of the event \\
\code{end.date} & Date of the last day of the event \\
\code{start.doy} & Day of the year of the first day of the event (1 = Jan. 1, etc.)\\
\code{start.month} & Month in which the event started (1 = January) \\
\code{days.above.abs.thresh.1} & Number of days in the event above a specified absolute threshold (default is the number of days in the event above 80\textsuperscript{o}F / 26.7\textsuperscript{o}C, but this can be changed with the \code{absolute\_thresholds} argument in \code{gen\_hw\_set}) \\
\code{days.above.abs.thresh.2} & Number of days in the event above a specified absolute threshold (default is the number of days in the event above 85\textsuperscript{o}F / 29.4\textsuperscript{o}C) \\ 
\code{days.above.abs.thresh.3} & Number of days in the event above a specified absolute threshold (default is the number of days in the event above 90\textsuperscript{o}F / 32.3\textsuperscript{o}C) \\
\code{days.above.abs.thresh.4} & Number of days in the event above a specified absolute threshold (default is the number of days in the event above 95\textsuperscript{o}F / 35.0\textsuperscript{o}C) \\
\code{days.above.99th} & Number of days in the event above the 99\textsuperscript{th} percentile of the variable for the location, using the period specified with the \code{referenceBoundaries} argument in \code{gen\_hw\_set} as a reference for determining these percentiles \\
\code{days.above.99.5th} & Number of days in the event above the 99.5\textsuperscript{th} percentile of the variable for the location, using the period specified with the \code{referenceBoundaries} argument in \code{gen\_hw\_set} as a reference for determining these percentiles \\
\code{first.in.year} & Whether the event was the first to occur in its calendar year in the location \\
\code{mean.var.quantile} & The percentile of the average variable value during the event compared to the location's year-round distribution of the variable, based on the variable distribution for the location during the period specified by the \code{referenceBoundaries} argument in \code{gen\_hw\_set} \\
\code{max.var.quantile} & The percentile of the maximum variable value during the event compared to the location's year-round distribution of the variable, based on the variable distribution for the location during the period specified by the \code{referenceBoundaries} argument in \code{gen\_hw\_set} \\
\code{min.var.quantile} & The percentile of the minimum variable value during the event compared to the location's year-round distribution of the variable, based on the variable distribution for the location during the period specified by the \code{referenceBoundaries} argument in \code{gen\_hw\_set} \\
\code{mean.seasonal.var} & The location's average seasonal value of the variable (by default, season is set to May--September, but this can be changed with the \code{seasonal\_months} argument in \code{gen\_hw\_set}), based on the variable values for the location during the years specified by the \code{referenceBoundaries} argument in \code{gen\_hw\_set} \\
\code{mean.yearround.var} & The location's average year-round value of the variable, based on the variable values for the location during the years specified by the \code{referenceBoundaries} argument in \code{gen\_hw\_set} \\
\bottomrule
\end{tabular}
\caption{Extreme event characteristics measured by the \code{gen\_hw\_set} function in the \pkg{futureheatwaves} package. The left column gives the name of each variable's column in the extreme event datasets created by the \code{gen\_hw\_set} function. When characterizing extreme events below a threshold, like cold spells, appropriate alternatives are given for some columns (e.g., \code{days.below.abs.thresh.1}, \code{days.below.1st}).}
\label{tab:hwcharacteristics}
\end{table}

\subsection{Example data}\label{example-data}

We have included data files in the package to serve as example files so
that users can try this package before applying it to their own
directory of climate projection files. This example data comes from two
climate models that are a part of CMIP5: (1) the model of the Beijing
Climate Center, China Meteorological Administration (BCC)
\citep{xin2013introduction} and (2) the National Center for Atmospheric
Research's (NCAR's) Community Climate System Model, version 4 (CCSM4)
\citep{gent2011community}. We include one ensemble member from BCC
(r1i1p1) and two from CCSM (r1i1p1 and r2i1p1).

To ensure that the size of this example data is reasonably small, we
have only included projection data for grid points from these climate
models that are near five U.S. east coast cities: New York, NY;
Philadelphia, PA; Newark, NJ; Baltimore, MD, and Providence, RI.
Further, to keep the file sizes reasonably small, the historical
projections range over the years 1990 to 1999, while the future
projections are limited to 2060 to 2079. Users' applications of this
package will likely use directories with many more climate model
ensemble members and more locations; however, the operation of the
package is the same for this smaller example application as it would be
for a much larger application.

Once the \pkg{futureheatwaves} package is installed and loaded, the user
can find the local location of these files using R's \code{system.file}
function. In the later sections of this article, we show how to use the
package functions with these example files as inputs.

When using this package with data other than this example data, the
climate model output data will require some pre-processing. Data will
need to be saved in a specific format, with files stored in a specific
directory structure. This preprocessing can be done using some of the R
packages described earlier for working with CMIP5 data in R. Full
details of the required file and directory structure are provided in the
package's ``futureheatwaves'' vignette, while tips on conducting this
processing starting from CMIP5 netCDF files are given in the
``starting\_from\_netcdf'' vignette.

\subsection{Basic example of identifying and characterizing extreme
events}\label{basic-example-of-identifying-and-characterizing-extreme-events}

Once climate model output files are set up as specified in the
``futureheatwaves'' package vignette, the package can process them to
identify and characterize heat waves in each ensemble member's
projection for each location using the \code{gen\_hw\_set} function. For
example, to process the example climate model output data included with
the package, the user can run:

\begin{Schunk}
\begin{Sinput}
projection_dir_location <- system.file("extdata/cmip5",
                                       package = "futureheatwaves")
city_file_location <- system.file("extdata/cities.csv",
                                  package = "futureheatwaves")

gen_hw_set(out = "example_results",
           dataFolder = projection_dir_location ,
           dataDirectories = list("historical" = c(1990, 1999),
                                        "rcp85" = c(2060, 2079)),
           citycsv = city_file_location,
           coordinateFilenames = "latitude_longitude_NorthAmerica_12mo.csv",
           tasFilenames = "tas_NorthAmerica_12mo.csv",
           timeFilenames = "time_NorthAmerica_12mo.csv")
\end{Sinput}
\end{Schunk}

This code first identifies and saves as objects the path names on the
user's computer of the example climate projections directory
(\code{projection\_dir\_location}) and the file of study locations
(\code{city\_file\_location}). The \code{gen\_hw\_set} function
processes this example input and creates a new directory,
\file{example\_results}, with files of identified and characterized heat
waves, in the user's current working directory. In this example code,
the processing is done using default values for the event definition,
years for which to generate the heat wave data sets, etc. How and why to
customize these choices are explained later in the text. Function
arguments (e.g., \code{dataDirectories}, \code{tasFilenames}) are used
to specify the format of the data and the directory structure.

Once the function has completed running, results will be written locally
to the directory specified by the \code{out} argument of
\code{gen\_hw\_set}. This directory will include files with some basic
information about the climate models and the closest grid points of each
climate model to each location, as well as a directory with files of
identified and classified extreme events for each ensemble member,
including all characteristics in Table \ref{tab:hwcharacteristics}.

\subsection{Customizing the extreme event
definition}\label{customizing-the-extreme-event-definition}

By default, the package identifies extreme events in climate model
output data using a specific definition for heat waves:

\begin{quote}
A \dfn{heat wave} is two or more days at or above a city-specific
threshold temperature, with the threshold determined as the
98\textsuperscript{th} percentile of year-round temperature in the city
during some reference period (by default, 1990--1999).
\end{quote}

\noindent However, the user can extensively customize the definition
used to identify extreme events. Users can write a custom R function
with either a different heat wave definition (see \citet{smith2013heat}
and \citet{kent2014heat} for listings of definitions used in scientific
studies) or with a definition appropriate for a different type of
extreme event (e.g., one or more days at or below 32\textsuperscript{o}F
/ 0\textsuperscript{o}C for frost day spells). For heat wave
identification, researchers might want to use a specific definition, for
example, because it matches the definition used by local health
officials to declare heat wave warnings or, in the case of health impact
assessments, to match with a definition used in an epidemiological
study. For other extreme events, a heat wave definition likely will not
be applicable and so a customized definition is necessary.

Three components of the extreme event definition can be easily
customized in the \code{gen\_hw\_set} function call, without creating a
new R function to use to identify heat waves. First, often users wish to
define an extreme event based on conditions that are rare in the study
location \citep{IPCCch1}. In this case, the percentile of the variable
of interest required for an extreme event can be changed using the
\code{probThreshold} option in \code{gen\_hw\_set}. For example, this
option could be used to identify heat waves based on the
99\textsuperscript{th} percentile, rather than the
98\textsuperscript{th} percentile, of a city's year-round temperature.
Second, the user can change the number of days used in the heat wave
definition using the \code{numDays} argument in the \code{gen\_hw\_set}
function (the default is \samp{numDays = 2}, corresponding to two or
more days). Combined, these two customization choices allow the user to
identify heat waves using many of the heat wave definitions used in
previous climate and health research-- for example, 9 of the 16 heat
wave definitions used in \citet{kent2014heat} (e.g., {[}some examples of
the 9 that could be fit{]}) could be fit using different combinations of
these two options for specifying threshold percentile and number of
days. Third, some extreme events like cold waves and frost day spells
are defined as a certain number of days below, rather than above, a
threshold. While the default is to identify events by searching for days
above a threshold, this behavior can be changed with the
\code{above\_threshold = FALSE} argument in the
\code{gen\_hw\_set function}.

Beyond these simpler options, the customization of the event definition
is even more extensive as one has the option of writing a custom R
function. This functionality allows the user to use heat wave
definitions that either require a number of days above an absolute
threshold (e.g., maximum temperature of \(\ge\) 95\textsuperscript{o}F
for \(\ge1\) day \citet{kent2014heat, tan2007heat}; minimum temperature
\(\le\) 0\textsuperscript{o}C for \(\ge1\) day for frost day spells) or
that require a combination of thresholds to be met (e.g., maximum daily
temperature one threshold every day of the heat wave and above a higher
threshold for a certain number of days;
\citet{kent2014heat, peng2011toward}). To use a customized event
definition, the user can write and load an R function that implements
the definition. The custom function is passed to the \code{gen\_hw\_set}
function using the \code{IDheatwavesFunction} argument. To work
correctly, this custom function must allow only specific inputs and
generate only specific outputs. Details about this required structure
are provided in the \pkg{futureheatwaves} package vignette. To increase
processing speed when identifying extreme events, we coded parts of the
default event definition function in C++ and synced it with R using the
\CRANpkg{Rcpp} package \citep{Rcpp}. Users should consider a similar
strategy for custom heat wave definitions, especially if processing a
large number of climate projection files.

Finally, extreme events are often based on conditions that are rare for
a specific location using location-specific relative thresholds. These
thresholds are often defined for climate impact studies based on a
variable's distribution at that location in present-day or historical
data. However, it can be interesting to explore trends in extreme events
under climate change if extreme events are identified based on variable
distributions during the projection period, or another future period. It
is possible to specify the time period to use when determining a
relative threshold for an event definition using the
\code{thresholdBoundaries} argument in the function \code{gen\_hw\_set}.

\subsection{Exploring extreme events}\label{exploring-extreme-events}

Once the user has created a directory of characterized event files for
each ensemble member (``Extreme event data sets'', Figure
\ref{fig:overview}), he or she can explore the results using the
\code{apply\_all\_models} function. This function allows the user to
apply custom R functions across all extreme event data frames created by
the \code{gen\_hw\_sets} call. The user can apply any R function that
follows certain standards in accepting input and returning output. Full
details on these standards are given in the \pkg{futureheatwaves}
package vignette.

As an example, if the user wanted to calculate the average temperature
of the heat waves identified for each ensemble member, he or she could
write a simple function:

\begin{verbatim}
average_mean_temp <- function(hw_datafr){
        out <- mean(hw_datafr$mean.var)
        return(out)
        }
\end{verbatim}

\noindent This function could then be applied across all extreme event
data sets output by \code{gen\_hw\_set} using the
\code{apply\_all\_models} function. For example, to apply this function
to all the example output results that come with the package, the user
could run:

\begin{Schunk}
\begin{Sinput}
out <- system.file("extdata/example_results", package = "futureheatwaves")
apply_all_models(out = out, FUN = average_mean_temp)
\end{Sinput}
\begin{Soutput}
#>   model ensemble    value
#> 1  bcc1        1 302.3745
#> 2  ccsm        1 302.4458
#> 3  ccsm        2 302.3428
\end{Soutput}
\end{Schunk}

This output gives the results (\code{value} column) of running the
custom function for each ensemble member of each climate model. Note
that the location of the directory with the heat wave data frames must
be specified using the \code{out} argument when calling
\code{apply\_all\_models}. Typically, this will be the directory path
for the directory specified with the \code{out} argument in
\code{gen\_hw\_set}.

Location-specific results can be generated using the
\code{city\_specific} argument in \code{apply\_all\_models}:

\begin{Schunk}
\begin{Sinput}
apply_all_models(out = out, FUN = average_mean_temp, city_specific = TRUE)
\end{Sinput}
\begin{Soutput}
#>    model ensemble city    value
#> 1   bcc1        1 balt 305.1816
#> 2   bcc1        1  nwk 300.3367
#> 3   bcc1        1   ny 300.3367
#> 4   bcc1        1 phil 305.1816
#> 5   bcc1        1 prov 298.0402
#> 6   ccsm        1 balt 303.1277
#> 7   ccsm        1  nwk 302.4053
#> 8   ccsm        1   ny 302.4053
#> 9   ccsm        1 phil 302.3425
#> 10  ccsm        1 prov 301.8895
#> 11  ccsm        2 balt 302.9373
#> 12  ccsm        2  nwk 302.2748
#> 13  ccsm        2   ny 302.2748
#> 14  ccsm        2 phil 302.2858
#> 15  ccsm        2 prov 301.9520
\end{Soutput}
\end{Schunk}

This output is structured as ``tidy" data \citep{wickham2014tidy},
allowing it to be used easily with the graphing package
\CRANpkg{ggplot2} \citep{ggplot2}.

The \code{apply\_all\_models} function can also be used to project the
health impacts of heat waves. A simple estimate of excess deaths
associated with heat wave risk in a community can be calculated as
\citep{peng2011toward}:

\begin{equation}
E_c = B_c * (RR - 1) * L_c 
\label{eq:healthimpacts}
\end{equation}

where:

\begin{itemize}
\item $E_c$ is the total number of excess deaths in community $c$;
\item $B_c$ is the baseline average daily mortality in community $c$; 
\item $RR$ is the relative risk of mortality associated with a heat wave day; and 
\item $L_c$ is the total number of heat wave days in community $c$ over the study period.
\end{itemize}

This impact assessment equation can be translated into a function that
merges each projection's heat wave data frame with a data frame of
community-specific baseline mortality rates (\(B_c\)), calculates
equation \ref{eq:healthimpacts} for each heat wave, and then sums up the
total excess deaths across all heat waves. In this example, we have set
the default relative risk for a heat wave as 1.032, based on the
estimated added heat wave relative risk in 107 U.S. communities for heat
waves defined as two or more days at or above a community's
98\textsuperscript{th} percentile temperature
\citep{anderson2009weather}.

\begin{verbatim}
excess_deaths <- function(hw_datafr, base_mortality, RR = 1.032){
        hw_datafr <- dplyr::left_join(hw_datafr, base_mortality,
                                      by = "city") %>%
                dplyr::mutate(excess_deaths = base_mort * (RR - 1) * length)
        out <- sum(hw_datafr$excess_deaths)
        out <- round(out)
        return(out)
}
\end{verbatim}

\noindent Once defined in R, this function can be applied across all
heat waves from all climate models' ensemble members, provided that you
have a data frame called \code{city\_mortality} with columns with each
community's identifier (\code{city}) and baseline mortality rate
(\code{base\_mort}), using the call (this example uses hypothetical
baseline mortality rates for the cities and requires the \CRANpkg{dplyr}
package \citep{dplyr}):

\begin{Schunk}
\begin{Sinput}
library(dplyr)
city_mortality <- data.frame(city = c("balt", "nwk", "ny", "phil", "prov"),
                             base_mort = c(3, 2.5, 35, 6, 2.5))
apply_all_models(out = out, FUN = excess_deaths, base_mortality = city_mortality)
\end{Sinput}
\begin{Soutput}
#>   model ensemble value
#> 1  bcc1        1   930
#> 2  ccsm        1   968
#> 3  ccsm        2   967
\end{Soutput}
\end{Schunk}

\noindent While the above impact analysis is simplistic, it provides an
example of how the \code{apply\_all\_models} function can be used in
conjunction with epidemiologic results to estimate the potential health
impacts of extreme events with climate change.

\section{Customizing dates for projections and reference
temperatures}\label{customizing-dates-for-projections-and-reference-temperatures}

Depending on the climate model output data used and the research
question, different projection periods may be of interest. For example,
one researcher may wish to explore extreme events in the 2050--2075
period, while another may be interested in 2090--2100. The projection
date range can be changed by specifying starting and ending years in the
\code{projectionBoundaries} argument of the \code{gen\_hw\_set}
function.

Further, some of the event characteristics (e.g.,
\code{mean.temp.quantile}, Table \ref{tab:hwcharacteristics}) are
calculated based on relative temperature, which is a measure of how the
value of the variable of interest during an extreme event compares to
the typical distribution of that variable at that location. These
characteristics are measured for each of the extreme events identified
by the \code{gen\_hw\_set} function by taking the absolute value of the
variable during the event (e.g., average temperature during the heat
wave is 90\textsuperscript{o}F ~32.2\textsuperscript{o}C) and comparing
it to the location's typical variable distribution. This process
generates relative measures of how intense the event is compared to what
is normal in that location (e.g., 90\textsuperscript{o}F
~32.2\textsuperscript{o}C is in the 99\textsuperscript{th} percentile of
year-round temperatures in the location).

These relative event characteristics will vary depending on whether you
calculate them based on a location's present-day variable distribution
or on the location's variable distribution in the future, since the
distributions of many relevant variables (e.g., temperature,
precipitation) are expected to change in many locations with climate
change. The package therefore allows the user to specify date ranges of
the temperature distributions to be used in calculating these relative
temperature metrics in each location, which can be done using the
\code{referenceBoundaries} option of \code{gen\_hw\_set}.

\subsection{Mapping grid points}\label{mapping-grid-points}

\begin{figure}
\begin{center}
\includegraphics[width = 0.9\textwidth]{ExampleLeaflet}
\end{center}
\caption{Snapshot of an example interactive map created using the \code{map\_grid\_leaflet} showing the locations of study cities and their matching climate model grid points for the BCC climate model example data included with \pkg{futureheatwaves}. The lines on the map connect each climate model grid point to the study location(s) for which that grid point was used. The interactive maps include pop-ups with city identiers; one is shown open in this snapshot as an example. From this map, you can see that the climate model grid point closest to New York City for this climate model is over the Atlantic Ocean.}
\label{fig:gridmap}
\end{figure}

It can be useful to explore the location of the climate model grid point
used to pull climate model output for each study location with a given
climate model. Therefore, the package has a function called
\texttt{map\_grid\_leaflet} that plots the locations of grid points used
for each location from each climate model. This function is built using
the htmlWidget \pkg{leaflet} package \citep{leaflet}. The following code
illustrates the use of this function with the example data to create
Figure \ref{fig:gridmap}, which plots the grid points used in the
example data from the BCC climate model in the example data:

\begin{Schunk}
\begin{Sinput}
out <- system.file("extdata/example_results", package = "futureheatwaves")
map_grid_leaflet(plot_model = "bcc1", out = out)
\end{Sinput}
\end{Schunk}

\subsection{Extensions}\label{extensions}

While this package was created to be used for research on extreme events
in climate change projections, it can be used more broadly. For example,
there are other episodes like wildfires and air pollution where it may
be interesting to identify extended periods of high exposures in
projection time series, and this package could be applied to gridded air
pollution model output to explore these exposures.

\section{Acknowledgements}\label{acknowledgements}

This work was supported by grants from the National Institute of
Environmental Health Sciences (R00ES022631), the National Science
Foundation (1331399), and the Colorado State University Vice President
for Research. Thank you to early testers of the package: Julia
Bromberek, Wande Benka-Coker, Josh Ferreri, Ryan Gan, Molly Gutilla,
Mike Lyons, Casey Quinn, Rachel Severson, and Meilin Yan.

\bibliography{Anderson}

\address{%
G. Brooke Anderson\\
Colorado State University\\
Department of Environmental \& Radiological Health Sciences\\ 1681 Campus Delivery\\ Fort Collins, Colorado 80523\\
}
\href{mailto:brooke.anderson@colostate.edu}{\nolinkurl{brooke.anderson@colostate.edu}}

\address{%
Colin Eason\\
Colorado State University\\
Department of Computer Science\\ 1873 Campus Delivery\\ Fort Collins, Colorado 80523\\
}
\href{mailto:aimesce@gmail.com}{\nolinkurl{aimesce@gmail.com}}

\address{%
Elizabeth A. Barnes\\
Colorado State University\\
Department of Atmospheric Science\\ 1371 Campus Delivery\\ Fort Collins, CO 80523\\
}
\href{mailto:eabarnes@atmos.colostate.edu}{\nolinkurl{eabarnes@atmos.colostate.edu}}