### 2017-06-01 paper_drafts/RJournalRevisions2/response_to_reviewers.txt
Create submission for R Journal

Thank you very much for the opportunity to revise and resubmit this manuscript. The
reviewers made some excellent suggestions, and we have modified the manuscript to
respond to these. In cases where suggested changes are beyond the scope of the described
package, we have added text to the manuscript to explain how such proposed approaches could
contribute as R programmers continue to evolve software solutions to working with climate model
output in R, and we have tried to make some connections to how R-based approaches to working with
large data continue to evolve in other topical areas with large, array-based data (e.g., flow
cytometry). In addition, we have revised and updated the package that is described by this
manuscript, both in response to reviewer suggestions and also to ensure functionality with recent
updates to `dplyr` and other key imported packages that have undergone major updates since the
original submission. This updated version of the package is available through GitHub
(https://www.github.com/geanders/futureheatwaves), and we will submit this updated package
version to CRAN as soon as we have satisfactorily responded to reviewers.

Comments from Reviewer #1

Comment: The authors answered quite satisfactory to the questions and also greatly
improved the manuscript and the package. I recommend that the article
should be published "as is" after solving a minor issue of the package: the
function 'map_grid_leaflet' does not work correctly because it requires a
local copy of the leaflet library.  Is it possible to put the http link of
leaflet instead of a local one so that function works for any connected
machine?

Response: The current version of the package imports the latest version of
`leaflet` from CRAN (version 1.1.0) and works with this version of leaflet,
rather than a local copy of the leaflet library. The function will now
work for any connected machine.

Comments from Reviewer #2

Comment: Climate science is fundamental to informing key societal decisions in the 21st
century. The Climate Intercomparison Project (currently in Phase 5) is a valuable
suite of Early system model outputs that continues to produce novel science. As
such, R packages to ingest and analyze these outputs could be quite valuable to the
community, particularly to field scientists and experimentalists who are not familiar
with the more idiosyncratic pipelines traditionally used by Earth system modelers. That
being said, this package does not address the most pressing barriers to data access and
interpretation, and I'm unclear that this manuscript describes a package that is
sufficiently novel to warrent publication.

Response: We appreciate the reviewer's point that the described package does not
resolve all the challenges of working with climate model output data in R.
However, we think that the package described offers an important step forward in making the
ensemble techniques that have become popular in climate science research more accessible to
researchers who are conducting climate impact studies but are not themselves climate scientists.
We have added a "Motivation" subsection to the manuscript section describing the `futureheatwaves`
package to clarify this point:

Added text: "We created the `futureheatwaves` package to aid in identifying, characterizing, and
exploring multi-day extreme events in daily climate model output data. While most of the discrete
tasks involved in identifying and characterizing multi-day extreme events are fairly
straightforward, the full process can be code-intensive, especially for multi-city studies and
studies that test sensitivity to how an event is defined or that incorporate different scenarios of
adaptation in the case of events defined using a threshold relative to community climate. Our aim
in developing this package was therefore to make the full process of identifying and characterizing
these extreme events much more convenient and so facilitate the use of multi-model, multi-ensemble
member analyses in climate impact studies conducted by non-climate scientists."

Further, we have also added a section to the end of the paper ("Future directions for working
with climate model output in R") that describes how future software development in R could
continue to tackle the challenges provided by climate data, to help put the package described in
the current manuscript in context within the evolving use of R for research involving climate
model output data.

Comment: This article correctly identifies memory use and data format as a key issue for the
usability of CMIP5 data, neither of which are addressed by this package. The user is required
to use `ncdf4` to preprocess the CMIP5 outputs and script their own sub-setting routines. This
sidesteps both the memory issue and data formatting. The comments on calendars being non-standard
in the CMIP5 models is, again, useful and accurately identified but not addressed systematically
in the package and left as an issue for the user to solve, albeit, with guidance. If parts of this
process were hard coded (for example, Model A has a 365-day calendar starting from 1800 and model
B has a leap-year calendar starting from 1900) this could be very useful to the community.

Response: We appreciate the reviewer's point that having information at hand on which calendar
a climate models uses might be useful to those working with climate model output in R. Such a
table, however, would be fairly complex if it included both the calendar and the reference date
from which days are counted for all CMIP5 data. This is because, for some
of the CMIP5 climate models, different calendars are used for different
experiments, and difference reference dates are also used for different experiments within climate
models.

To provide some guidance to R users, however, we have added a link to the manuscript for a table
provided in the Frequently Asked Questions section of the Earth System CoG website. This table
gives the calendar (although not all starting reference dates) for each of the CMIP5 models,
including notes of exceptions for certain experiments within some models. Here is the sentence
we added to the manuscript with this link:

Added text: "While most CMIP5 climate models use the same calendar for all experiments, a few do not; a full table of the calendars used for each climate model and experiment, pulled from netCDF metadata, is available at https://www.earthsystemcog.org/projects/cog/faq_data."

In terms of the challenges presented by the large file sizes of climate
model output, we do recommend some pre-processing of data files to subset files down
to the geographic area and time range required for a study (e.g., North America for a national
study of the United States) which would help reduce data size prior to processing climate model
output through the described package. As the reviewer notes, we offer suggestions and guidance for
conducting this pre-processing in R, both through text in the manuscript. We offer additional
guidance through a vignette provided with the package described in the manuscript
(https://cran.r-project.org/web/packages/futureheatwaves/vignettes/starting_from_netcdf.html).
Even after pre-processing the files prior to using the `futureheatwaves` package, many research
applications will still involve working with very large files. For example, a national study of the
United States may still use files with daily temperatures for many decades at a large number of
grid points, and an analysis that uses multiple climate models, some with multiple ensemble
members, would still need to process a number of very large files. The `futureheatwaves` package
addresses these large file sizes by inputting and processing data from each ensemble member of each
climate model separately, avoiding the memory problems created by trying to load all data from all
ensemble members into memory at once and processing the data using `apply`, `map`, `plyr`, or
similar tools applied across all data at once.

We acknowledge, however, that while this approach is successful in terms of allowing the
identification and characterization of sustained extreme events in R, future software development
could continue refining and optimizing the handling of climate model data in R. We have added a
section to the end of the manuscript on future directions for climate model output data in R and
have included comments on how such future work could continue to tackle the challenges introduced
by the large file sizes of climate model output data:

Added text: "Another important path for development could be through approaches that allow
researchers to take advantage of the statistical tools offered by R while maintaining large climate
model output files in a netCDF format. For example, Goncalves and coauthors recently described a
"round table" approach of connecting as-needed data access from netCDF climate data files through
to functionality available in R and CDO through the intermediary of a MonetDB database system
(Goncalves et al., 2015). In other topical areas, R programmers are also improving the efficiency
of working with data in large netCDF files through approaches that avoid loading all data
in-memory. For example, the Bioconductor package `ncdfFlow` enables R users to conduct analyses of
hundreds of large flow cytometry output files through the creation and use of an `ncdfFlowSet`
class that stores the data in an HDF5 format rather than in-memory (Jiang et al., 2017;
Finak et al., 2014). Such an approach could be promising for future R software development for
working with large climate model files."

Comment: Some acknowledgement of NCL or CDO for merging, sub-setting, averaging, and otherwise
manipulating netCDF files is warranted in this manuscript. These processes are MUCH faster in
these specialty languages and then done via R and not to mention this when dealing with CMIP5
data is remiss. One idea, although this completely changes the focus of the package, would be
to create a wrapper for these speciality languages in R to leverage that speed.

Response: Thank you for this suggestion. In the revised version of the manuscript, we have
added references to both software tools, as well as text describing the availability of
NCL and CDO for initial processing of climate model output files, including both the computation speed advantages and the disadvantage of needing to learn the syntax of a new language:

"CMIP5 data are available in the netCDF file format. Free specialty software exists to work with
climate model output files in this format, including a collection of command line tools developed
by the Max Planck Institute called Climate Model Operators (CDO) (Schulzweida, 2017) and an
interpreted language developed by the National Center for Atmospheric Research (NCAR) called the
NCAR Command Language (NCL) (UCAR/NCAR Computational and Information SystemsLaboratory, 2017).
Although such software can be used to quickly process netCDF climate model output files, they
require learning a new language syntax, and so for R users may not be worth the computational speed
gain compared to alternative solutions that can be scripted in the R language."

Further, we have added a section to the end of the paper, in response to this and other comments
from this reviewer, called "Future directions for working with climate model output in R". In this
section, we have added some text outlining the reviewer's idea that wrappers for these two
speciality languages could be an interesting next step to pursue as R software for working with
climate model output data evolves, including references to existing libraries with wrappers in
other languages (Python and Ruby):

Added text: "One important step in future development of R software to work with climate model output could be the development of R wrappers for some of the existing command line tools available
through the Climate Data Operators (CDO) software (Schulzweida, 2017). Libraries already exist for
Python and Ruby that allow the functionality of CDO tools to be used within these scripting
languages (available from the Max Planck Institute at
https://code.zmaw.de/projects/cdo/wiki/Cdo%7Brbpy%7D). While one R package (`ncdf4.helpers`;
Bronaugh, 2014) already provides R wrappers for a few CDO operators, such functionality could be
extended through future R software to capture more of the full functionality of the CDO toolkit."

Comment: The analysis that this package does is fairly straightforward using `plyr` or other
common R packages / functions. The main function described in this article is the identifying
and summing ranges of events. If a user made it far enough to do the pre-processing of the CMIP5
output data I doubt this would be a computational barrier to further analysis. If authors could
show notable speed or memory requirement improvement then this could be interesting.

Response: As noted in a previous response, the advantage of this package is to improve
convenience for researchers identifying and characterizing sustained, multi-day events in
daily climate model output, especially for analyses that incorporate multiple climate
models and ensemble members. We have added a paragraph to the manuscript making this
motivation for the described package clearer:

Added text: "We created the `futureheatwaves` package to aid in identifying,
characterizing, and exploring multi-day extreme events in daily climate model output data.
While most of the discrete tasks involved in identifying and characterizing multi-day extreme
events are fairly straightforward, the full process can be code-intensive, especially for
multi-city studies and studies that test sensitivity to how an event is defined or that incorporate
different scenarios of adaptation in the case of events defined using a threshold relative to
community climate. Our aim in developing this package was therefore to make the full process of
identifying and characterizing these extreme events much more convenient and so facilitate the use
of multi-model, multi-ensemble member analyses in climate impact studies conducted by non-climate
scientists."

Comment: Readability of the article is excellent.

Response: Thank you. We have tried to maintain the same level of readability through our revisions
in response to reviewers' comments.